---
title: "Homework 8"
author: "Kaleb Luna | kl36927"
date: "2025-04-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

GitHub Link: <https://github.com/kalebjoel/Homework-8-.git>

<P style="page-break-before: always">

\newpage

# Problem 1: Regression Warm up 

```{r echo = FALSE, results = FALSE}
library(ggplot2)
creatnineData <- read.csv("creatinine.csv")

```

```{r echo = FALSE, results = FALSE, message = FALSE}
ggplot(creatnineData, aes(x = age, y = creatclear)) + 
  geom_point() + 
  geom_smooth(method = 'lm', se = FALSE) + 
  labs(title = "Creatinine Clearance vs. Age", 
       x = "Age (years)",
       y = "Creatinine Clearance")

lm1 = lm(creatclear ~ age, data = creatnineData)
coef(lm1) |> round(2) 


# y = -x 
```

## Part A

The expected creatinine clearance rate for a 55-year-old is approximately **113.71 mL/minute.** This expected rate was determined by obtaining the regression equation below from the visualization of the linear regression line above.

**Equation:**

-   ŷ = 147.81 - 0.62x

## Part B 

```{r echo = FALSE, results = FALSE}
coef(lm1)
```

For every year of age, creatinine clearance **decreases** by approximately **-0.62 mL/minute.**

-   I obtained this value by fitting a linear regression model using the lm() function, using creatnine clearance as the response variable and age as the explanatory variable. Then, I utilized the coef() function to obtain the slope of the regression line. \

## Part C

```{r echo = FALSE, results = FALSE}
predict(lm1, newdata = data.frame(age= c(50,60)))
```

Based on the regression model **(ŷ = 147.81 - 0.62x)**, the predicted creatinine clearance rate of a 40-year old is 116.82 mL/minute, and the predicted creatinine clearance rate for a 60-year old is 110.62 mL/minute. Comparing the actual values with the predicted, we see:

-   40-year-old: 135 - 116.82 = 18.18 mL/min above expected

-   60-year-old: 112 - 110.62 = 1.38 mL/min above expected

The creatinine clearance rate is healthier in the 40-year-old with a rate of 135 compared to the 60-year-old.

<P style="page-break-before: always">

\newpage

# Problem 2: Modeling Disease Growth

## Part 1 

```{r echo = FALSE, warning = FALSE, message = FALSE, results = FALSE}
covidData <- read.csv("covid.csv")

italy <- subset(covidData, country == "Italy")
modelItaly <- lm(log(deaths) ~ days_since_first_death, data = italy)
coef(modelItaly)
doublingItaly <- log(2) / 0.183218

spain <- subset(covidData, country == "Spain")
modelSpain <- lm(log(deaths) ~ days_since_first_death, data = spain)
coef(modelSpain)
doublingSpain <- log(2) / 0.2762447

```

```{r echo = FALSE, results = FALSE}
set.seed(1)
n <- nrow(italy)
nboot <- 1000

italy_slopes <- rep(NA, nboot)
italy_doubling <- rep(NA, nboot)

for (i in 1:nboot) {
  samp <- italy[sample(1:n, replace = TRUE), ]
  fit <- lm(log(deaths) ~ days_since_first_death, data = samp)
  slope <- coef(fit)[2]
  italy_slopes[i] <- slope
  italy_doubling[i] <- log(2) / slope
}

quantile(italy_slopes, c(0.025, 0.975))     
quantile(italy_doubling, c(0.025, 0.975))    

```

The estimated growth rate and doubling time for Italy is 0.183 and 3.8 days, respectively. The 95% bootstrapped confidence intervals for both measurements are (0.159, 0.209) and (3.323, 4.367), respectively.

## Part 2

```{r echo = FALSE, message = FALSE, results = FALSE}
set.seed(2)  
n <- nrow(spain)

spain_slopes <- rep(NA, nboot)
spain_doubling <- rep(NA, nboot)

for (i in 1:nboot) {
  samp <- spain[sample(1:n, replace = TRUE), ]
  fit <- lm(log(deaths) ~ days_since_first_death, data = samp)
  slope <- coef(fit)[2]
  spain_slopes[i] <- slope
  spain_doubling[i] <- log(2) / slope
}

quantile(spain_slopes, c(0.025, 0.975))     
quantile(spain_doubling, c(0.025, 0.975))    

```

The estimated growth rate and doubling time for Spain is 0.276 and 2.5 days, respectively. The 95% bootstrapped confidence intervals for both measurements are (0.234, 0.320) and (2.168, 2.958), respectively.

## Part 3

```{r echo = FALSE}

ggplot(covidData, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line() + 
  labs(title = "Daily COVID-19 Deaths Over Time", 
       x = "Days Since First Death", 
       y = "Number of Deaths")

```

<P style="page-break-before: always">

\newpage

# Problem 3: Price Elasticity of Demand 

```{r echo = FALSE, results= FALSE}
milkData <- read.csv("milk.csv")
milkModel <- lm(log(sales) ~ log(price), data = milkData)
summary(milkModel) 

set.seed(3)
n <- nrow(milkData)
nboot <- 1000
betas <- rep(NA, nboot)

for (i in 1:nboot) {
  samp <- milkData[sample(1:n, replace = TRUE), ]
  fit <- lm(log(sales) ~ log(price), data = samp)
  betas[i] <- coef(fit)[2]
}

quantile(betas, c(0.025, 0.975))  # 95% CI

```

The estimated price elasticity of demand for milk is approximately **-1.619,** meaning that the demand for milk decreases as the price increases. The estimated 95% bootstrapped confidence interval is **(-1.784, -1.460).**

To estimate the price elasticity of demand for milk, I fit a linear regression model using log(sales) and log(prices) based on the power-model equation, Q. = K \* P\^(Beta), from which I obtained the price elasticity estimate. Then, I used the bootstrapping technique to get a 95% confidence interval for the price elasticity.
